Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-08-12 Manuscript/journal_correspondence/ERL_responseletter.Rmd

\blankline

Dear Editors:

\blankline

Please consider this revised version of article ERL-111271, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_. We thank the two referees for their reviews of this manuscript, and have fully addressed each of their comments in detail below. Our point-by-point responses immediately follow each referee comment, and we include line numbers of the revised manuscript in each response to identify where we made changes to the text.

@@

## Referee 1, comments to the author:

**The manuscript describes an impressive effort projecting ecosystem carbon balance in the Hawaiian Islands under climate change and land use change scenarios. This assessment of future regional terrestrial carbon dynamics can provide a support to local planning of climate mitigation strategies through ecosystem carbon sequestration. The manuscript is well written and organized. However, more clarification and justification are needed for methodology and assumptions about how future NPP and land cover change is modelled. And also, more details are needed such as maps illustrating how projected change in carbon flux (e.g., NPP and NBP) and carbon pools (live biomass, soil organic soil) is spatially distributed. I have provided detailed comments below.**

As suggested by this reviewer, we have added clarification and justification about how future NPP and land cover change were modeled, as well as new figures illustrating the spatial distribution of current and projected future carbon fluxes. Our responses to individual comments below indicate where we made these changes.

**L109, why need to assign each cell a state type? What is consideration for it?**  The state of each cell tracks a range of information about that cell, including land cover class, moisture zone, island, age, and time since transition. This set of information is used to define the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur. We have added text and references to the first paragraph in the Methods section that defines the state of each cell (lines 101-102).

**L114, transitions between ‘state type’? Is it ‘land cover classes’? According to the definition of state type as unique combination of moisture zones, islands and land cover classes at L110, is there transition from dry forest and mesic forest in the same island?**  
We agree this was unclear. We have added text to the first paragraph of the Methods section that defines "transition" (line 103) and we have re-written this paragraph to make clear that transition pathways define which land cover classes can be converted to other land cover classes (lines 121-128).

**L123-126, high land use scenario samples historical rate between median and maximum, but why not low land use scenario sample rates between minimum and median? What is the justification that low land use scenario will have rate lower than historical records?**  
Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, reducing over time the total land area under active cultivation in the Hawaiian Islands. We consider this to be a reasonable possibility under a low land use scenario, especially given the volatility in agricultural land area change experienced in the Hawaiian Islands over the recent historical period (1992-2010). We have added text that explains why we bounded agricultural land area change rates between zero and the minimum 1992-2010 historical rates in the low land use scenario (lines 134 - 137).

**L126, missing description for urbanization rate for high land use scenario.**  
Transition rates for urbanization in the high land use scenario were sampled from a uniform distribution bounded by the median and maximum historical rate, as stated on lines 121-123 of the original manuscript. We have re-worded this description to make it more clear that it includes urbanization (lines 131 - 134).

**L133, what transitions can increase forest area such as low land use scenario shown in figure S5? Does tree plantation refer to reforestation or afforestation? If afforestation, what is the source land cover class for this plantation?** 
Forest area increases in the low land use scenario because of agricultural contraction (i.e., abandonment of previously cultivated land). We have added text to the supplementary explaining that the increase in forest, grassland, and shrubland area in the low land use scenario between 2020 and 2040 is largely due to agricultural contraction over the same time period (supplementary lines xxx - xxx). Tree plantation does not refer to reforestation or afforestation. We did not include expansion of tree plantations as a transition type, and so there is no land cover class that can be converted to tree plantation in either land use scenario. Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture, which partially explains the increase in grassland area in both land use scenarios (supplemental Figure 7). We have added text to the Methods section that better describes tree plantation harvest rates and transitions (lines 143-152). The source of all land cover classes, including tree plantations, is the land cover class map depicted in Figure 1, derived from Jacobi et al. (2017).

**L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**  
We have added text defining the historical harvest rotation rate as between 5-7 years, as well as more detail on how tree plantation area is projected to change under both land use scenarios (lines 143 - 152).

**L137, will wildfire take place only in forests? If so, what is the resulting land cover class? Would it still recover back to forest after a certain time?**  
Wildland fire in Hawaiʻi primarily ignites in non-native grasslands and shrublands, but can sometimes spread into forests. We have added text and references that describe where wildland fire occurs in the Hawaiian Islands (lines 163 - 168).

@@

**L140, is state type land cover classes? Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**  
We agree this was a confusing use of terminology. We have changed "state type" to "land cover class" here (line xxx) and elsewhere throughout the manuscript when it refers only to land cover classification. We now use the term "state type" only when referring to the unique combination of island, moisture zone, and land cover class.

**L158, moisture zone is included in state type as defined at line L110, isn’t it?**  
This was a confusing use of terminology. We have omitted "moisture zone" here and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition as the unique combination of island, moisture zone, and land cover class (lines 115-117).

@@

**L162, how litter and dead wood pools were initialized?**  
We have added text describing how litter and dead wood pools were initialized using values from the IBIS DGVM (lines xxx-xxx).

**L161, what is 30-year climate normal come from? Does it have consistent temperature and rainfall distribution with RCP projections during 2010 and 2020? Or is there any bias between 30-year climate normal and RCP projections?**  
The contemporary 30-year climate normals for rainfall and temperature come from the references cited (Giambelluca et al. 2013, 2014), as depicted in supplementary Figure 3. We have added text to the supplementary that better describes statistical downscaling of CMIP5 temperature and rainfall projections and how they compare to contemporary climate data (supplementary lines xxx-xxx).

**L163, figure 2 shows total live biomass in 2010 is about 68 Tg C, but total forest aboveground biomass is estimated as 36 Tg C in Asner et al 2016 which is used for initialization here. Where is the additional 32 Tg C from? What is ratio of aboveground biomass to total live biomass in IBIS?**  
Asner et al. (2016) only quantifies aboveground forest live biomass. The additional 32 Tg C comes from the sum of belowground forest live biomass and total live biomass in grasslands, shrublands, and agriculture. We have added text and citations describing how we used a power function applied to the Asner et al. (2016) abovegrould forest live biomass values to estimate forest belowground live biomass and, subsequently, estimate total forest live biomass, which was used to calibrate the IBIS DGVM (lines 184-186).

**L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
We have added text (lines 188-189) listing the range of GPP values from the Kimball et al. (2017) calibration dataset and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (supplementary table 1).

**L166-168, is similar multiplier also applied to living biomass? If not, will this single NPP multiplier break the equilibrium between NPP and biomass? For example, for a simulation cell, NPP and biomass should reach dynamic equilibrium in IBIS simulation, a given NPP could only support carbon turnover of a certain number of trees. External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**  
No, there is no set of spatial multipliers applied to living biomass. We have added text and figures to the supplementary that better describes how we derived and applied the NPP multipliers (supplementary lines xxx - xxx, supplementary figures 5 & 6). We have also added text to the manuscript describing how the LUCAS initialization of carbon stocks and fluxes allowed for these stocks and fluxes to reach a dynamic equilibrium across the landscape of the Hawaiian Islands (lines 204-209).

**Line 173, why assume Rh as ratio of NPP? Is the ratio constant over time and across climate change scenarios? Why not to estimate Rh based on soil moisture, temperature and soil carbon stocks? Because of lack of soil temperature in RCP 4.5 and 8.5? Need more justification for this flux ratio method.** 
We have added text and citations justifying the fact that plant productivity is the primary factor driving variation in ecosystem carbon loss at annual timescales and that the vast majority of annual carbon loss from soils was fixed by photosynthesis within that year (lines 199-202). Although we agree that gridded annual maps of projected future soil moisture and soil temperature could potentially improve predictions of ecosystem carbon flux in our model, these data products do not exist for the Hawaiian Islands.

**L177-179, are the growth and decay multipliers varied at annual basis? How are multipliers derived for time periods CMIP5 projections do not cover? Such as period between 2020-2049? It is very necessary to include the distribution of these multipliers, are they enhancing or weakening NPP over time at two RCPs?** 
Yes, the growth and decay multipliers vary annually. We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines 210-215, supplementary lines xxx-xxx, supplementary Figures 5 & 6).

@@

**L182, the empirical model used here to relate NPP to temperature and rainfall is core to projections of carbon balance. It is necessary to present this empirical relationship in the paper, such as what the optimal temperature of NPP is and how it changes with rainfall. For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**  
We have added a supplemental figure displaying the spatial distribution of projected future NPP multipliers based on future climate projections (supplemental Figure 6). We have also added text to the supplementary that better describes how we employ the empirical NPP equations to calculate spatial NPP anomalies for both contemporary and projected future NPP (supplemental lines xxx).

**L184, what is the temporal coverage and resolution of these climate projections? Do them only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described in L179?**  
Yes, climate projections only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described on line 179 of the original manuscript. We have added text to the supplementary Climate section that provides more detail on climate projections (supplementary lines xxx).

**L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**  
Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.1 Tg y^-1^ in 2010 to 7.95 Tg y^-1^ in 2020), and the ranges of statewide annual NPP estimates broadly overlapped for each of these years (2011 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^). This slight decline in modeled NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change. We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the Results section (lines 274-276).

### 2021-08-12 Manuscript/HI_Model_ms.Rmd
updated manuscript files and response letter

```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in total ecosystem carbon storage (a) and individual carbon stocks (b) for the seven main Hawaiian Islands. Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
knitr::include_graphics("./fig_images/fig2_Cstocks.png")
```
Net primary production (NPP) for the seven main Hawaiian Islands was ~8.1 Tg y^-1^ at the beginning of the simulation period in 2010 (supplementary figure 5) with land use change driving an approximate 2% decline during the contemporary period (2010-2020). NPP continued to decline throughout the rest of the simulation period (2020-2100) across all four scenarios, but this long-term decline was driven primarily by climate change and to a lesser extent by land use change (Fig. 3). The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figures 4 and 6) and sustained losses of forest and shrubland land area in the high land use scenario (supplemental figure 7). In contrast, climate change led to increased heterotrophic respiration (R~h~) over time, such that more intense warming under RCP 8.5 (supplemental figure 4) resulted in R~h~ being 3% higher on average by 2100 than under RCP 4.5 (figure 3). Heterotrophic respiration declined substantially over time in the high land use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 7), similar to trends in NPP. Transition-triggered carbon fluxes to the atmosphere from land use, land use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~ 0.4 Tg y^-1^ in the high land use scenario and ~ 0.2 Tg y^-1^ in the low land use scenario (figure 3). Uncertainty around transition-triggered carbon fluxes were higher in the high land use scenario, driven primarily by greater variability in wildland fire probabilities.

>>>>

```{r fig.align = 'center', out.width = "99%", fig.cap = "Projected changes in total ecosystem carbon storage (a) and individual carbon stocks (b) for the seven main Hawaiian Islands. Solid lines indicate the mean of 30 Monte Carlo realizations for each scenario, with shaded areas indicating the minimum and maximum range of Monte Carlo realizations."}
knitr::include_graphics("./fig_images/fig2_Cstocks.png")
```
Net primary production (NPP) for the seven main Hawaiian Islands was ~8.1 Tg y^-1^ at the beginning of the simulation period in 2010 (supplementary figure 5) with land use change driving an approximate 2% decline during the contemporary period (2010-2020). NPP continued to decline throughout the rest of the simulation period (2020-2100) across all four scenarios, but this long-term decline was driven primarily by climate change and to a lesser extent by land use change (Fig. 3). The combination of high radiative forcing (RCP 8.5) and high land use change led to the steepest decline in NPP over time, driven by intense long-term drying on the leeward sides of islands under RCP 8.5 (supplemental figures 4 and 6) and sustained losses of forest and shrubland land area in the high land use scenario (supplemental figure 7). In contrast, more intense warming under RCP 8.5 (supplemental figure 4) led to an increase in heterotrophic respiration (R~h~) in the latter half of the 21st century, such that R~h~ was 3% higher on average by 2100 in the RCP 8.5 radiative forcing scenario than in the RCP 4.5 radiative forcing scenario (figure 3). Heterotrophic respiration declined substantially over time in the high land use scenario (figure 3) because of long-term reductions in forest and shrubland land area (supplemental figure 7), similar to trends in NPP. Transition-triggered carbon fluxes to the atmosphere from land use, land use change, and wildfire were largely independent of changes in climate, stabilizing by mid-century at an average of ~ 0.4 Tg y^-1^ in the high land use scenario and ~ 0.2 Tg y^-1^ in the low land use scenario (figure 3). Uncertainty around transition-triggered carbon fluxes were higher in the high land use scenario, driven primarily by greater variability in wildland fire probabilities.
